---
title: "bootstrap"
author: "Leonor Rui"
date: "2024-11-14"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(p8105.datasets)

set.seed(1)
```

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Bootsrapping!

Make up some data

```{r}
n_samp = 250

sim_df_constant = 
  tibble(
    x = rnorm(n_samp, 1, 1),
    error = rnorm(n_samp, 0, 1),
    y = 2 + 3 * x + error
  )
 
## error based on x
sim_df_nonconstant = 
  sim_df_constant |>
  mutate(error = error * 0.75 * x,
         y = 2 + 3 * x + error)
```

Look at these

```{r}
sim_df_constant |>
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  stat_smooth(method = "lm")

sim_df_nonconstant |>
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  stat_smooth(method = "lm")
```

Look at regression results

```{r}
sim_df_nonconstant |>
  lm(y ~ x, data = _) |>
  broom::tidy() |>
  knitr::kable(digits = 3)

sim_df_constant |>
  lm(y ~ x, data = _) |>
  broom::tidy() |>
  knitr::kable(digits = 3)
```

## Draw a bootsrap sample

```{r}
boot_sample = function(df) {
  
  boot_df = sample_frac(df, replace = TRUE)|>
    arrange(x)
  
  return(boot_df)
}
```

Let's try running this

```{r}
## Certain points show up multiple times
sim_df_nonconstant |>
  boot_sample() |>
  ggplot(aes(x = x, y = y)) +
  geom_point(alpha = 0.5) +
  stat_smooth(method = "lm")
```

Do this as part of analysis

```{r}
sim_df_nonconstant |>
  boot_sample() |>
  lm(y ~ x, data = _) |>
  broom::tidy() |>
  knitr::kable(digits = 3)
```

## Bootsrap A LOT

(Construct a simulated distribution of slope)

analyze 1000 bootsrap samples, take the standard deviation of the estimates (Standard Error)

Because the assumption of normal error term is violated, the boostrap SEs are not very similar to the original ones

```{r}
boot_straps = 
  tibble(
    strap_number = 1:1000
  ) |>
  mutate(
    ## run the function for i times, i = strap_number
    strap_sample = map(strap_number, \(i) boot_sample(df = sim_df_nonconstant)),
    models = map(strap_sample, \(df) lm(y ~ x, data = df)),
    results = map(models, broom::tidy)
  )

bootstrap_results = 
  boot_straps |>
  select(strap_number, results) |>
  unnest(results) |>
  group_by(term) |>
  summarise(
    boot_se = sd(estimate)
  )
```

If the assumption of normal error is satisfied, the bootsrap SEs will be almost the same with the original ones

```{r}
boot_straps = 
  tibble(
    strap_number = 1:1000
  ) |>
  mutate(
    ## run the function for i times, i = strap_number
    strap_sample = map(strap_number, \(i) boot_sample(df = sim_df_constant)),
    models = map(strap_sample, \(df) lm(y ~ x, data = df)),
    results = map(models, broom::tidy)
  )

bootstrap_results = 
  boot_straps |>
  select(strap_number, results) |>
  unnest(results) |>
  group_by(term) |>
  summarise(
    boot_se = sd(estimate)
  )
```

## Do this all using modelr

```{r}
boot_straps =
  sim_df_nonconstant |>
  modelr::bootstrap(1000) |>
  mutate(
    strap = map(strap, as_tibble),
    models = map(strap, \(df) lm(y ~ x, data = df)),
    results = map(models, broom::tidy)
  ) |>
  select(.id, results) |>
  unnest(results)
```

what you want to report

```{r}
boot_straps |>
  group_by(term) |>
  summarise(
    boot_est = mean(estimate),
    boot_se = sd(estimate),
    boot_cl_ll = quantile(estimate, .025),
    boot_cl_ul = quantile(estimate, .975)
  )
```

## AirBNB dataset

```{r}
data("nyc_airbnb")

manhattan_df = nyc_airbnb |>
  mutate(stars = review_scores_location / 2) |>
  rename(borough = neighbourhood_group, neighborhood = neighbourhood) |>
  filter(borough == "Manhattan") |>
  select(price, stars, room_type) |>
  drop_na()
```

Plot the data

```{r}
manhattan_df |>
  ggplot(aes(x = stars, y = price)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)
```

fit a regression

```{r}
manhattan_df |>
  lm(price ~ stars + room_type, data = _) |>
  broom::tidy() |>
  knitr::kable(digits = 3)
```

Bootstrap  for better inference

```{r}
manhattan_df |>
  modelr::bootstrap(1000) |>
  mutate(
    strap = map(strap, as_tibble),
    models = map(strap, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(models, broom::tidy)
  ) |>
  select(.id, results) |>
  unnest(results) |>
  filter(term == "stars") |>
  ggplot(aes(estimate)) +
  geom_density()
```

Find the true distribution of the estimate of price 










