---
title: "cross_validation"
author: "Leonor Rui"
date: "2024-11-12"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
library(SemiPar)

set.seed(1)
```

Look at LIDAR data

```{r}
data("lidar")

lidar_df = lidar |>
  as_tibble() |>
  mutate(id = row_number())
```

```{r}
lidar_df |>
  ggplot(aes(x = range, y = logratio)) +
  geom_point()
```

## Try to do CV

We'll compare 3 models -- one linear, one smooty, one wiggly

Construct training and testing df

```{r}
train_df = sample_frac(lidar_df, size = .8)
test_df = anti_join(lidar_df, train_df, by = "id")
```

```{r}
ggplot(train_df, aes(x = range, y = logratio)) +
  geom_point() +
  geom_point(data = test_df, color = "red")
```

Fit three models

```{r}
linear_mod = lm(logratio ~ range, data = train_df)

smooth_mod = gam(logratio ~ s(range), data = train_df)

## Set more turns
wiggly_mod = gam(logratio ~ s(range, k = 30), sp = 10e-6, data = train_df)
```

Look at fits

```{r}
train_df |>
  add_predictions(linear_mod) |>
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
  geom_line(aes(y = pred), color = "red")

train_df |>
  add_predictions(smooth_mod) |>
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
  geom_line(aes(y = pred), color = "red") +
  geom_point(data = test_df, color = "blue")

train_df |>
  add_predictions(wiggly_mod) |>
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
  geom_line(aes(y = pred), color = "red") +
  geom_point(data = test_df, color = "blue")
```

Compare these numerically using RMSE

```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)
```

## Repeat the train/test split

```{r}
cv_df = 
  crossv_mc(lidar_df, 100) |>
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

```{r}
cv_df |>
  pull(train) |>
  nth(3)
```

Fit models, extract RMSEs

```{r}
## Fit linear model to each training dataset
## For each linear model, fit it to the test dataset, then get the rmse
cv_res_df = 
  cv_df |>
  mutate(
    linear_mod = map(train, \(x) lm (logratio ~ range, data = x)), 
    smooth_mod = map(train, \(x) gam(logratio ~ s(range), data = train_df)), 
    wiggly_mod = map(train, \(x) gam(logratio ~ s(range, k = 30), sp = 10e-6, data = train_df))
  ) |>
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, rmse), 
    rmse_smooth = map2_dbl(smooth_mod, test, rmse), 
    rmse_wiggly = map2_dbl(wiggly_mod, test, rmse)
  )
```

Look at RMSE distribution

```{r}
cv_res_df |>
  select(starts_with("rmse")) |>
  pivot_longer(
    everything(), 
    names_to = "model",
    values_to = "rmse", 
    names_prefix = "rmse_"
  ) |>
  ggplot(aes(x = model, y = rmse)) +
  geom_violin()
```

## Nepalese children df

```{r}
child_df = read_csv("data/nepalese_children.csv") |>
  mutate(
    ## True = 1, False = 0
    weight_ch7 = (weight > 7) * (weight - 7)
  )
```

Look at data

```{r}
child_df |>
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = 0.5)
```

Fit some models

```{r}
linear_mod2 = lm(armc ~ weight, data = child_df)
pwl_mod = lm(armc ~ weight + weight_ch7, data = child_df)
smooth_mod2 = gam(armc ~ s(weight), data = child_df)
```

Look at the models

```{r}
child_df |>
  add_predictions(linear_mod2) |>
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red")

child_df |>
  add_predictions(smooth_mod2) |>
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red")

child_df |>
  add_predictions(pwl_mod) |>
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red")
```

CV to select models

```{r}
cv_df2 = 
  crossv_mc(child_df, 100) |>
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

Apply models and extract RMSE

```{r}
cv_res_df2 = 
  cv_df2 |>
  mutate(
    linear_mod = map(train, \(x) lm(armc ~ weight, data = child_df)),
    pwl_mod = map(train, \(x) lm(armc ~ weight + weight_ch7, data = child_df)), 
    smooth_mod = map(train, \(x) gam(armc ~ s(weight), data = child_df))
  ) |>
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, rmse),
    rmse_pwl = map2_dbl(pwl_mod, test, rmse),
    rmse_smooth =  map2_dbl(smooth_mod, test, rmse)
  )
```

```{r}
cv_res_df2 |>
  select(starts_with("rmse")) |>
  pivot_longer(
    everything(), 
    names_to = "model",
    values_to = "rmse", 
    names_prefix = "rmse_"
  ) |>
  ggplot(aes(x = model, y = rmse)) +
  geom_violin()
```









