---
title: "linear_models"
author: "Leonor Rui"
date: "2024-11-07"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

set.seed(1)
```

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


Do some data cleaning

```{r}
data("nyc_airbnb")

nyc_airbnb = nyc_airbnb |>
  mutate(stars = review_scores_location / 2) |>
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) |>
  filter(borough != "Staten Island") |>
  select(price, stars, borough, neighborhood, room_type) |>
  ## change into order by frequency
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type)
  )
```

## Fit some models

Start relatively simple

```{r}
fit = lm(price ~ stars, data = nyc_airbnb)

summary(fit)

names(summary(fit))

coef(fit)

## format the output into a dataframe
broom::tidy(fit)

fit |>
  broom::tidy() |>
  select(term, estimate, p.value)

fit |>
  broom::glance()
```

Try a bit more complex

R put each level of a categorical variable to be its own variable

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)

fit |>
  broom::tidy() |>
  select(term, estimate, p.value) |>
  mutate(
    term = str_replace(term, "borough", "Borough:")
  ) |>
  knitr::kable(digits = 3)
```

## Some diagnostics

backtrack to some EDA

```{r}
nyc_airbnb |>
  ggplot(aes(x = stars, y = price)) +
  geom_point() +
  stat_smooth(method = "lm")
```


Most diagnostics use residuals

```{r}
## Add residuals to the dataframe based on this model fit
modelr::add_residuals(nyc_airbnb, fit) |>
  ggplot(aes(x = resid)) +
  geom_histogram()

modelr::add_residuals(nyc_airbnb, fit) |>
  ggplot(aes(x = borough, y = resid)) +
  geom_violin() +
  ylim(-200, 500)
```

Residuals against stars

```{r}
modelr::add_residuals(nyc_airbnb, fit) |>
  ggplot(aes(x = stars, y = resid)) +
  geom_point()
```

Residuals against fitted values

```{r}
nyc_airbnb |>
  modelr::add_residuals(fit) |>
  modelr::add_predictions(fit) |>
  ggplot(aes(x = pred, y = resid)) +
  geom_point()
```

## Hypothesis Testing

- For single coefficients, look at the table

```{r}
fit |>
  broom::tidy()

fit_null = lm(price ~ stars + borough, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
```

Look at both

- When adding more than one coefficient, use "ANOVA"

```{r}
fit_null |>
  broom::tidy()

fit_alt |>
  broom::tidy()


## Only used for nested models 
anova(fit_null, fit_alt) |>
  broom::tidy()
```

## Do effects differ across boroughs

First use a lot of interactions

```{r}
nyc_airbnb |>
  lm(price ~ stars * borough + room_type * borough, data = _) |>
  broom::tidy()
```

Could just fit separate models 

```{r}
nyc_airbnb |>
  filter(borough == "Manhattan") |>
  lm(price ~ stars + room_type, data = _) |>
  broom::tidy()

nyc_airbnb |>
  filter(borough == "Brooklyn") |>
  lm(price ~ stars + room_type, data = _) |>
  broom::tidy()
```

Get fancy and use list columns

```{r}
nyc_airbnb |>
  nest(data = -borough) |>
  mutate(
    model = map(data, \(x) lm(price ~ stars + room_type, data = x)),
    results = map(model, broom::tidy)
  ) |>
  select(borough, results) |>
  unnest(results) |>
  select(borough, term, estimate) |>
  pivot_wider(
    names_from = term,
    values_from = estimate
  )
```

What about room type across manhattan neighborhoods?

```{r}
nyc_airbnb |>
  filter(
    borough == "Manhattan",
    neighborhood == "Chinatown") |>
  lm(price ~ stars + room_type, data = _) |>
  broom::tidy()

nyc_airbnb |>
  filter(
    borough == "Manhattan",
    neighborhood == "Chelsea") |>
  lm(price ~ stars + room_type, data = _) |>
  broom::tidy()
```

Let's do this for all neighborhoods

```{r}
manhattan_neighborhood_fit_df = 
  nyc_airbnb |>
    filter (borough == "Manhattan") |>
    nest( data = -(borough:neighborhood)) |>
    mutate(
      model = map(data, \(x) lm(price ~ stars + room_type, data = x)), 
      results = map(model, broom::tidy)
    ) |>
    select (neighborhood, results) |>
    unnest(results)
```

Look at the effect of room type 

```{r}
manhattan_neighborhood_fit_df |>
  filter(str_detect(term, "room_type")) |>
  ggplot(aes(x = term, y = estimate)) +
  geom_boxplot()
```

```{r}
manhattan_neighborhood_fit_df |>
  filter(str_detect(term, "room_type")) |>
  filter(estimate > 0)

nyc_airbnb |>
  filter(
    neighborhood == "NoHo",
    room_type == "Shared room"
  )
```

## Binary Outcomes

```{r}
baltimore_df = 
  read_csv("data/homicide-data.csv") |> 
  filter(city == "Baltimore") |> 
  mutate(
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")) |> 
  select(resolved, victim_age, victim_race, victim_sex)
```

```{r}
fit_logistic = 
  baltimore_df |> 
  glm(resolved ~ victim_age + victim_race + victim_sex, data = _, family = binomial()) 
```

```{r}
fit_logistic |> 
  broom::tidy() |> 
  mutate(OR = exp(estimate)) |>
  select(term, log_OR = estimate, OR, p.value) |> 
  knitr::kable(digits = 3)
```

```{r}
baltimore_df |> 
  modelr::add_predictions(fit_logistic) |> 
  mutate(fitted_prob = boot::inv.logit(pred))
```









